---
title: "ST 558 - Project 1"
output: 
  rmarkdown::github_document: 
    toc: true
    toc_depth: 2
author: "Christine Marie Heubusch"
date: "June 12, 2020"
---

# Understanding JSON Data

## What is JSON data?

[JavaScript Object Notation (JSON)](https://www.json.org/json-en.html) is a format of storing data that is "lightweight" (that is, tends to be a small file size) and integrates well with most programming languages, though it is technically a subset of JavaScript. These properties make it a favorite format for application programming interfaces (APIs) and is often used to ["serialize and transfer data over a network connection"](https://www.w3resource.com/JSON/introduction.php). In JSON, [data can be stored in a variety of different ways](https://www.youtube.com/watch?v=iiADhChRriM&list=TLPQMTAwNjIwMjBzt1zaeUU28g&index=2), including strings, numbers, booleans (with the true/false operators), arrays, and null. While objects are wrapped with a left and right curly brace ({}), arrays are enclosed in brackets ([]). Objects contain [key-value pairs](https://www.w3schools.com/js/js_json_objects.asp), with the key being a string, a value being one of the aformed JSON data types, and a colon separating the two; each key-value pair is separated with a comma. Values may be nested in a JSON data object, which makes it both convenient and easy to read.

## Packages Used for Reading JSON Data
Discuss the possible packages/functions that are available for reading JSON data into R. (There are three major packages for JSON data.) Choose one and explain why you’ve chosen it.

Three R packages can be utilized for working with JSON data: 

* `rjson`  
* `RJSONIO`  
* `jsonlite`  

All three packages offer options for converting to/from JSON. `RJSONIO` was originally introduced as a [faster alternative](https://www.rdocumentation.org/packages/RJSONIO/versions/1.3-1.4) to `rjson`, given concerns about the latter's slower speeds, though there are some indications that `rjson` may now actaully be faster in certain circumstances. I have selected `jsonlite` - which debuted in 2014 and originally [started as a branch of `RJSONIO`](https://cran.r-project.org/web/packages/jsonlite/index.html#:~:text=In%20addition%20to%20converting%20JSON,data%20in%20systems%20and%20applications.) - as my package of choice. Documentation on the package is extensive, as evidenced by the package's thorough [vignette](https://cran.r-project.org/web/packages/jsonlite/jsonlite.pdf); and a wide variety of useful functions and options are within. Interestingly, only `RJSONIO` and `jsonlite` [give control over vector simplification](https://rstudio-pubs-static.s3.amazonaws.com/31702_9c22e3d1a0c44968a4a1f9656f1800ab.html), an important consideration when parsing JSON data.

# Contacting the NHL API
```{r Loading Packages, include=FALSE}
library(httr)
library(jsonlite) 
library(knitr)
library(tidyverse)
```

When I first tried converting the data from JSON, *"data."* appeared at the beginning of each column name; ostensibly, this points to the object within which the array of data was stored in JSON. However, thinking this a bit cumbersome for usage in later code, per Subhankar's recommendation, I played around with using [`str_remove`](https://www.rdocumentation.org/packages/stringr/versions/1.4.0/topics/str_remove) to remove the string, and found it to be useful for keeping the column names short and more enjoyable to work with. 

## Creating a Function for Accessing Franchise Data
```{r Creating Function for Accessing Franchise, message=FALSE}
api_function_franchise <- function(x) {
  base_url <- "https://records.nhl.com/site/api" 
  x <- "franchise"
  full_url <- paste0(base_url, "/", x) 
  franchise_get <- GET(full_url) 
  franchise_txt <- content(franchise_get, "text") 
  franchise_json <- fromJSON(franchise_txt, flatten=TRUE, simplifyDataFrame=TRUE)
  franchise_df <- as.data.frame(franchise_json)
  colnames(franchise_df) <- str_remove(colnames(franchise_df), "data.") 
  return(franchise_df)
}
franchise_data <- api_function_franchise(x)
#view(franchise_data)
head(franchise_data)
```

## Creating a Function for Accessing Franchise Team Totals
```{r Creating Function for Accessing Franchise Team Totals, message=FALSE}
api_function_totals <- function(x) {
  base_url <- "https://records.nhl.com/site/api" 
  x <- "franchise-team-totals"
  full_url <- paste0(base_url, "/", x) 
  team_totals_get <- GET(full_url) 
  team_totals_txt <- content(team_totals_get, "text") 
  team_totals_json <- fromJSON(team_totals_txt, flatten=TRUE)
  team_totals_df <- as.data.frame(team_totals_json)  
  colnames(team_totals_df) <- str_remove(colnames(team_totals_df), "data.") 
  return(team_totals_df)
}
team_totals_data <- api_function_totals(x)
#view(team_totals_data)
head(team_totals_data)
```

## Creating a Function for Accessing Season Records
```{r Creating Function for Accessing Season Records, message=FALSE}
api_function_season <- function(franchiseID) {
  base_url <- "https://records.nhl.com/site/api" 
  full_url <- paste0(base_url, "/", "franchise-season-records?cayenneExp=franchiseId=", franchiseID) 
  season_get <- GET(full_url) 
  season_txt <- content(season_get, "text") 
  season_json <- fromJSON(season_txt, flatten=TRUE)
  season_df <- as.data.frame(season_json)
  colnames(season_df) <- str_remove(colnames(season_df), "data.") 
  return(season_df)
}
seasonCaps <- api_function_season(24) #Testing the function with Washington Capitals data, since I'm originally from the DC area!
#view(seasonCaps)
```

## Creating a Function for Accessing Goalie Records
```{r Creating Function for Accessing Goalie Records, message=FALSE}
api_function_goalie <- function(franchiseID) {
  base_url <- "https://records.nhl.com/site/api" 
  full_url <- paste0(base_url, "/", "franchise-season-records?cayenneExp=franchiseId=", franchiseID) 
  goalie_get <- GET(full_url) 
  goalie_txt <- content(goalie_get, "text") 
  goalie_json <- fromJSON(goalie_txt, flatten=TRUE)
  goalie_df <- as.data.frame(goalie_json)
  colnames(goalie_df) <- str_remove(colnames(goalie_df), "data.")
  return(goalie_df)
}
goalieCanes<- api_function_goalie(26) #Testing the function with Carolina Hurricanes data, to honor my new home. 
#view(goalieCanes)
```

## Creating a Function for Accessing Skater Records
```{r Creating Function for Accessing Skater Records, message=FALSE}
api_function_skater <- function(franchiseID) {
  base_url <- "https://records.nhl.com/site/api" 
  full_url <- paste0(base_url, "/", "franchise-season-records?cayenneExp=franchiseId=", franchiseID) 
  skater_get <- GET(full_url) 
  skater_txt <- content(skater_get, "text") 
  skater_json <- fromJSON(skater_txt, flatten=TRUE)
  skater_df <- as.data.frame(skater_json)
  colnames(skater_df) <- str_remove(colnames(skater_df), "data.")
  return(skater_df)
}
skaterPenguins <- api_function_skater(17) #Testing the function - sure, I suppose I'll look at some Penguins data too! 
#view(skaterPenguins)
```

# Exploratory Data Analysis

## Is a Team With More Recorded Shutouts More Likely to Still be an Active Franchise? 

In a [shutout game](https://en.wikipedia.org/wiki/Shutout), a hockey team's defense manages to prevent the other team from scoring any points throughout the entirety of the game - surely, a sign of strong performance! To begin, I decided to created a categorical variable called **shutoutRange**, creating ranges for total number of shutouts and using `ifelse` to assign each value to the appropriate level of the new variable. I then created a contingency table to get a quick glimpse of how many active (and inactive) franchises fell within each category. Not too surprisingly, none of the teams that had a history of over a 100 shutouts were inactive. To visualize this data, I created a barplot, filling the bars with color based upon whether or not the franchise is active.

```{r Creating shutoutRange Variable}
team_totals_data <- mutate(team_totals_data, shutoutRange = 
                             ifelse(shutouts %in% 0:50, "0-50",
                                ifelse(shutouts %in% 51:100, "51-100",
                                ifelse(shutouts %in% 101:150, "101-150", 
                                ifelse(shutouts >150, "151+", "")))))
team_totals_data$shutoutRange <- factor(team_totals_data$shutoutRange, levels=c("0-50", "51-100", "101-150","151+")) #Changed these values to factors so that they could be easily ordered for the barplot below. 
```

```{r Contingency Table of Active or Non-Active Franchise vs. Shutouts}
shutoutTable <- table(team_totals_data$activeFranchise, team_totals_data$shutoutRange)
kable(shutoutTable, caption="Table of Active or Non-Active Franchise vs. Total Number of Shutouts")
```

```{r Creating Bar Plot of Number of Shutouts, with Active Franchise}
shutoutGraph<- ggplot(data=team_totals_data, aes(x=shutoutRange))
shutoutGraph + geom_bar(aes(fill=as.factor(activeFranchise))) + 
  labs(x="Total Number of Shutouts", title="Bar Plot of Number of Shutouts") + 
  scale_fill_discrete(name="Active Franchise?", labels=c("No","Yes")) + 
  coord_flip()
```

## Are Teams more Likely to Win at Home than Away? 

We've heard of home-"field" advantage - let's see if the same can be said of the rink.I began by creating another two variables: 

* **roadWinLossRatio**, AKA total number of wins on the road divided by total number of losses on the road  
* **homeWinLossRaio**, AKA total number of losses at home divided by total number of wins on the road  

*For this purposes of this variable, I did not include ties.*

```{r Create roadWinLossRatio and homeWinLossRatio variables}
team_totals_data <- mutate(team_totals_data, roadWinLossRatio = roadWins / roadLosses)
team_totals_data <- mutate(team_totals_data, homeWinLossRatio = homeWins / homeLosses)

team_totals_data %>%
  select(teamName, roadWinLossRatio, homeWinLossRatio) %>%
  arrange(teamName) %>%
  head()
```

```{r Calculating roadWinToLossRatio Summary Statistics}
roadWLR <- summary(team_totals_data$roadWinLossRatio)
roadWLR
```

```{r Calculating homeWinToLossRatio Summary Statistics}
homeWLR <- summary(team_totals_data$homeWinLossRatio)
homeWLR
```

Looking at the mean and median win-to-loss ratios for on the road (or "away") and home, we see that both the home mean and median ratios are nearly double for home games - perhaps indicating that a team is more likely to win when it's playing on its own ice. We can then look at boxplots of these ratios, as a visual representation of these summary statistics.  

```{r Creating Boxplot of roadWinToLossRatio}
penalty_plot <- ggplot(team_totals_data, aes(x=roadWinLossRatio,y=""))
penalty_plot + 
  geom_boxplot(fill="#cc0000") +
  ggtitle("Boxplot of Ratio of Road Wins to Road Losses") +
  labs(x="Ratio of Road Wins, Divided by Road Losses", y="")
```

```{r Creating Boxplot of home WinLossRatio, warning=FALSE}
penalty_plot <- ggplot(team_totals_data, aes(x=homeWinLossRatio,y=""))
penalty_plot + 
  geom_boxplot(fill="#cc0000") +
  ggtitle("Boxplot of Ratio of Home Wins to Home Losses") +
  labs(x="Ratio of Home Wins, Divided by Home Losses", y="")
```

```{r Creating Histogram of Road Wins to Losses Ratio}
roadWinsLossesHist <- ggplot(team_totals_data, aes(x=roadWinLossRatio))
roadWinsLossesHist +
  geom_histogram(aes(y=..density..), binwidth=0.25) + 
  ggtitle("Histogram of Road Wins : Losses Ratio")
```

```{r Creating Histogram of Home Wins to Losses Ratio, warning=FALSE}
homeWinsLossesHist <- ggplot(team_totals_data, aes(x=homeWinLossRatio))
homeWinsLossesHist +
  geom_histogram(aes(y=..density..), binwidth=0.25) + 
  ggtitle("Histogram of Home Wins : Losses Ratio")
```

## Do Aggressive Teams Tend to Win More? 

Do teams that spend more time in the penalty box tend to win more? To help explore this question, I created two new variables from the Franchise Team Totals dataset (that is, R object team_totals_data). 

Here, I used the `mutate` function from the `dplyr` package to create a new variable called **penaltyGamesRatio** - the total number of minutes spent in the penalty box, divided by the total number of games played. This variable is intended to display the average amount of time per game a team's players spend in penalty - perhaps a larger number is indicative of a more "aggressive" playing style. 

```{r Create penaltyGamesRatio Variable}
team_totals_data <- mutate(team_totals_data, penaltyGamesRatio = penaltyMinutes / gamesPlayed)

team_totals_data %>%
  arrange(desc(penaltyGamesRatio)) %>%
  select(teamName, penaltyGamesRatio) %>%
  head()
```

I then calculated an **overall** "win-to-loss" ratio, by dividing total number of wins by total number of losses. My hope here was to create a standardized measure, controlling for the amount of time that a team had been in operation, so that new teams and old teams could be compared equitably. Upon printing the first six rows of the data, we see that the Vegas Golden Knights - a team that just began in 2017 - has the highest win-to-loss ratio. 
```{r Calculate Overall Win-Loss Ratio}
team_totals_data <- mutate(team_totals_data, winLossRatio = wins / losses)
team_totals_data %>%
  select (teamName, winLossRatio) %>%
  arrange(desc(winLossRatio)) %>%
  head()
```

I then created another variable: **MoreWinsOrLosses**, using `ifelse`. If a team won more than it lost (that is, its winLossRatio > 1), then the value of MoreWinsOrLosses was designated as "More Wins". Otherwise, it was assigned a value of "More Losses". *(For the purposes of this project, I designated the scenario of an equal number of wins and losses as "More Losses".)*
```{r Create Categorical Variable MoreWinsOrLosses}
team_totals_data <- mutate(team_totals_data, moreWinsOrLosses = ifelse(winLossRatio > 1, "More Wins", "More Losses"))
team_totals_data %>%
  select(teamName, winLossRatio, moreWinsOrLosses) %>%
  arrange(teamName) %>%
  head(10)
```

```{r Contingency Table of Franchise vs. Shutouts}
shutoutTable <- table(team_totals_data$activeFranchise, team_totals_data$shutoutRange)
kable(shutoutTable, caption="Table of Active or Non-Active Franchise vs. Total Number of Shutouts")
```

```{r Contingency Table of More Wins or Losses vs. Shutout Range}
kable(table(team_totals_data$moreWinsOrLosses, team_totals_data$shutoutRange), caption="Contingency Table of More Wins or Losses and Shutout Range")
```

Looking at the resulting boxplot *(below)*, we can see that the median amount of time spent in the penalty box by a "winning" team is greater than that of a "losing" team. Yet interestingly, the winning team range is much, much shorter than that of a losing team, and two losing teams prove to be outliers, with one team spending an average of *over 30 minutes a game* in the penalty box. 

```{r Creating Boxplot of Average Time Spent in Penalty Box}
penalty_plot <- ggplot(team_totals_data, aes(x=penaltyGamesRatio,y=moreWinsOrLosses))
penalty_plot + 
  geom_boxplot(fill="#cc0000") +
  ggtitle("Boxplot of Average Time Spent in Penalty Box") +
  labs (x="Ratio of Total Minutes in Penalty Box to Total Number of Games Played", y="'Winning' or 'Losing' Team?")
```

Looking at the scatterplot, does not appear to be a straight-line relationship between the amount of time that a team spends in the penalty box and its Win-Loss ratio. Looking at the distribution of the points, it seems that there may be a "sweet spot" - perhaps indicating that making certain illegal or more aggressive plays that could be risky, but be rewarding. 

```{r Creating Scatterplot of the Penalty-Games Ratio and Win-Loss Ratio, message=FALSE}
ratiosScatter <- ggplot(team_totals_data, aes(x=penaltyGamesRatio, y=winLossRatio))
ratiosScatter +
  geom_point() +
  geom_smooth() + 
  ggtitle("Scatterplot of Average Time Spent in Penalty Box vs. Win-Loss Ratio") + 
  labs(x="Average Time Spent in Penalty Box", y="Win-Loss Ratio")
```

## Do Newer Teams Tend to be More Aggressive? 

For this brief analysis, I filtered the dataset twice: once for years after or equal to 2000 (season stored as 20002001), then for years before 2000. I then took summary statistics for each of the ratios of Penalty to Games. In this case, we see that the median and mean for teams before 2000 is actually higher than that of teams that began in or after 2000, perhaps indicating that newer teams are actually less likely to be in the penalty box.

```{r Getting Penalty Summary Statistics for Teams Started in or After 2000}
post2000Penalties<- team_totals_data %>% 
  filter(firstSeasonId >= 20002000) %>% 
  select(teamName, penaltyGamesRatio, firstSeasonId) %>% 
  arrange(desc(penaltyGamesRatio))
summary(post2000Penalties$penaltyGamesRatio)
```

```{r Getting Penalty Summary Statistics for Teams Started Before 2000}
pre2000Penalties <- team_totals_data %>% 
  filter(firstSeasonId < 20002000) %>% 
  select(teamName, penaltyGamesRatio, firstSeasonId) %>% 
  arrange(desc(penaltyGamesRatio))
summary(pre2000Penalties$penaltyGamesRatio)
```

## How are Weak Goalies Affecting Teams' Performances? 

The [Goals Against Average](https://en.wikipedia.org/wiki/Goals_against_average) indicates "the number of goals a goaltender allows per 60 minutes of playing time" - that is, per one game. The column goalsAgainst in the dataset is a cumulative measure, indicating the total number of Goals Against for the team since its inception. We can average this as a "per game" value by dividing the goalsAgainst by the total number of games played. After creating that new variable, I then plotted this value against the total Win-Loss ratio. Of course, there appears to be a negative correlation between the two values: the larger the number of goals that a team's goalies have let in, the lower their win-loss ratio. I colored the values on the scatterplot by whether or not a team is still an active franchise - indeed, the greater the number of goals are let in, the lower the number of wins... and seemingly, the less likely they are to still be an active franchise. The active franchises appear to have a much lower Goals Against Ratio, and a much higher Win-Loss ratio. 

```{r Creating Scatterplot of Goals Against vs. Losses, message=FALSE}

team_totals_data$activeFranchise <- as.logical(team_totals_data$activeFranchise)
team_totals_data <- mutate(team_totals_data, goalsAgainstRatio = goalsAgainst / gamesPlayed)

goalsAgainstLossesPlot <- ggplot(team_totals_data, aes(x=goalsAgainstRatio, y=winLossRatio, color=activeFranchise))
goalsAgainstLossesPlot + 
  geom_point() + 
  geom_smooth(method=lm) +
  ggtitle("Scatterplot of Goals Against vs. Win-Loss Ratio") +
  labs(x="Goals Against Ratio", y="Win-Loss Ratio")
```




